---
author: "Jasmine Lantaca"
---

In this lab, we will be using the `dplyr` package to explore student evaluations
of teaching data. 

**You are expected to use functions from `dplyr` to do your data manipulation!**

# Part 1: GitHub Workflow

Now that you have the Lab 5 repository cloned, you need to make sure you can 
successfully push to GitHub. To do this you need to:

-   Open the `lab-5-student.qmd` file (in the lower right hand corner). 
-   Change the `author` line at the top of the document (in the YAML) to your
name. 
-   Save your file either by clicking on the blue floppy disk or with a shortcut
(command / control + s). 
-   Click the "Git" tab in upper right pane
-   Check the "Staged" box for the `lab-5-student.qmd` file (the file you
changed)
-   Click "Commit"
-   In the box that opens, type a message in "Commit message", such as
"Added my name".
-   Click "Commit".
-   Click the green "Push" button to send your local changes to GitHub.

RStudio will display something like:

```         
>>> /usr/bin/git push origin HEAD:refs/heads/main
To https://github.com/atheobold/introduction-to-quarto-allison-theobold.git
   3a2171f..6d58539  HEAD -> main
```

Now you are ready to go! Remember, as you are going through the lab I would 
strongly recommend rendering your HTML and committing your after **every** 
question! 


# Part 2: Some Words of Advice

Part of learning to program is learning from a variety of resources. Thus, I
expect you will use resources that you find on the internet. There is, however,
an important balance between copying someone else's code and *using their code to learn*.  

Therefore, if you use external resources, I want to know about it.

-   If you used Google, you are expected to "inform" me of any resources you
used by **pasting the link to the resource in a code comment next to where you used that resource**.

-   If you used ChatGPT, you are expected to "inform" me of the assistance you
received by (1) indicating somewhere in the problem that you used ChatGPT 
(e.g., below the question prompt or as a code comment), and (2) downloading and
including the `.txt` file containing your **entire** conversation with ChatGPT.

Additionally, you are permitted and encouraged to work with your peers as you
complete lab assignments, but **you are expected to do your own work**. Copying
from each other is cheating, and letting people copy from you is also cheating.
Please don't do either of those things.

## Setting Up Your Code Chunks

-   The first chunk of this Quarto document should be used to *declare your libraries* (probably only `tidyverse` for now).
-   The second chunk of your Quarto document should be to *load in your data*.

## Save Regularly, Render Often

-   Be sure to **save** your work regularly.
-   Be sure to **render** your file every so often, to check for errors and make
sure it looks nice.
    -   Make sure your Quarto document does not contain `View(dataset)` or `install.packages("package")`, both of these will prevent rendering.
    -   Check your Quarto document for occasions when you looked at the data by
    typing the name of the data frame. Leaving these in means the whole dataset
    will print out and this looks unprofessional. **Remove these!**
    -   If all else fails, you can set your execution options to `error: true`,
    which will allow the file to render even if errors are present.

# Part 3: Let's Start Working with the Data!

## The Data

The `teacher_evals` dataset contains student evaluations of reaching (SET)
collected from students at a University in Poland. There are SET surveys from 
students in all fields and all levels of study offered by the university.

The SET questionnaire that every student at this university completes is as
follows:

> Evaluation survey of the teaching staff of University of Poland. Please
> complete the following evaluation form, which aims to assess the lecturer’s
> performance. Only one answer should be indicated for each question. The
> answers are coded in the following way: 5 - I strongly agree; 4 - I agree;
> 3 - Neutral; 2 - I don’t agree; 1 - I strongly don’t agree.
>
> Question 1: I learned a lot during the course.
>
> Question 2: I think that the knowledge acquired during the course is very
> useful.
>
> Question 3: The professor used activities to make the class more engaging.
>
> Question 4: If it was possible, I would enroll for a course conducted by this
> lecturer again.
>
> Question 5: The classes started on time.
>
> Question 6: The lecturer always used time efficiently.
>
> Question 7: The lecturer delivered the class content in an understandable and
> efficient way.
>
> Question 8: The lecturer was available when we had doubts.
>
> Question 9. The lecturer treated all students equally regardless of their
> race, background and ethnicity.

These data are from the end of the winter semester of the 2020-2021 academic
year. In the period of data collection, all university classes were entirely
online amid the COVID-19 pandemic. While expected learning outcomes were not
changed, the online mode of study could have affected grading policies and could
have implications for data.

**Average SET scores** were combined with many other variables, including:

1.  **characteristics of the teacher** (degree, seniority, gender, SET scores in
the past 6 semesters).
2.  **characteristics of the course** (time of day, day of the week, course
type, course breadth, class duration, class size).
3.  **percentage of students providing SET feedback.**
4.  **course grades** (mean, standard deviation, percentage failed for the
current course and previous 6 semesters).

This rich dataset allows us to **investigate many of the biases in student evaluations of teaching** that have been reported in the literature and to formulate new
hypotheses.

Before tackling the problems below, study the description of each variable
included in the `teacher_evals_codebook.pdf`.

**1. Load the appropriate R packages for your analysis.**

```{r}
#| label: setup
# code chunk for loading packages
library(tidyverse)
```

**2. Load in the `teacher_evals` data.** 

```{r}
#| label: load-data
# code chunk for importing the data

evals <- read_csv("Data/Data-Raw/teacher_evals.csv")
```

### Data Inspection + Summary

**3. Provide a brief overview (~4 sentences) of the dataset.**

```{r}
#| label: explore-data
# you may want to use code to answer this question

glimpse(evals)
dim(evals)
```
> There is a total of 8015 rows of data with 22 columns. The data is collected from students at Universirty of Poland from all sorts of courses and studies offered by the university whose rating the teaching staff of based on 9 question with answers from 1 to 5. Other data information includes the time of day with day of the week and class duration of the course the teacher teach, the gender of teacher, the degree and senority of the teacher, and data fo the students like average grade/standard deviation, and percentage of failed students. 

**4. What is the unit of observation (i.e. a single row in the dataset) identified by?**

```{r}
#| label: row-identification
# you may want to use code to answer this question

summary(evals)
```

> In a single observation, the data consist of 22 elements: course id, teacher id, the type of question answered, the total number of participants for teacher/course, proportion of survey response to the all, evaluation average of teacher, grade average of student in teacher class of both the past semesters and current semester, the standard deviation of student grade in teacher class for both current semester and past semesters, teacher's degree, senority, gender, course information (time of day, day of the week, duration of class), and percentage of student's failed in both past and current semester.

**5. Use _one_ `dplyr` pipeline to clean the data by:**

- **renaming the `gender` variable `sex`**
- **removing all courses with fewer than 10 respondents**
- **changing data types in whichever way you see fit (e.g., is the instructor ID really a numeric data type?)**
- **only keeping the columns we will use -- `course_id`, `teacher_id`, `question_no`, `no_participants`, `resp_share`, `SET_score_avg`, `percent_failed_cur`, `academic_degree`, `seniority`, and `sex`**

**Assign your cleaned data to a new variable named `teacher_evals_clean` –- use these data going forward. Save the data as `teacher_evals_clean.csv` in the `data-clean` folder.**

```{r}
#| label: data-cleaning
# code chunk for Q4

evals_clean <- evals |>
  rename(sex = gender) |>
  filter(no_participants > 10) |>
  mutate(across(c(teacher_id, question_no), as.factor)) |>
  select(c(course_id, teacher_id, question_no, no_participants, resp_share, SET_score_avg, percent_failed_cur, academic_degree, seniority, sex))
  

glimpse(evals_clean)

write_csv(evals_clean, "Data/Data-Clean/teacher_evals_clean.csv")
```

**6. How many unique instructors and unique courses are present in the cleaned dataset?**

```{r}
#| label: unique-courses
# code chunk for Q5


n_distinct(evals_clean$course_id) #number of unique courses
n_distinct(evals_clean$teacher_id) # number of teachers
```

> There are 921 unique courses while 294 teachers.

**7. One teacher-course combination has some missing values, coded as `NA`. Which instructor has these missing values? Which course? What variable are the missing values in?**

```{r}
#| label: uncovering-missing-values
# code chunk for Q6

evals_clean |>
  filter(if_any(.col = everything(), is.na))

```

> The teacher-course combination that has missing value is course PAB3SE004PA and teacher	56347. The variable that is missing is the percentage failed in the current semester.

**8. What are the demographics of the instructors in this study? Investigate the variables `academic_degree`, `seniority`, and `sex` and summarize your findings in ~3 complete sentences.**

```{r}
#| label: exploring-demographics-of-instructors
# code chunk for Q7

evals_clean |>
  select(sex, seniority, academic_degree) |>
  count(sex)

evals_clean |>
  select(sex, seniority, academic_degree) |>
  count(seniority) |>
  arrange(desc(n))

evals_clean |>
  select(sex, seniority, academic_degree) |>
  count(academic_degree)

```

> Looking at the demographics of the instructors, there are more males (3464) than female (3199) professors/teaching staffs. Majority of the professors have been working in the university at least 2 years. the second closest majority senority is 6 years, and third cloest majority is 8 years; this data shows that majority of teaching staffs are people with decent years of teaching in the university. In terms of academic degree, majority of teachers hold a doctorate degree compared to the rest.

**9. Each course seems to have used a different subset of the nine evaluation questions. How many teacher-course combinations asked all nine questions?**

```{r}
#| label: teacher-course-asked-every-question
# code chunk for Q8

evals_clean |>
  count(course_id, teacher_id) |> 
  filter(n == 9)

```

> There are 48 teacher-course combinations that asked all nine questions.

## Rate my Professor

**10. Which instructors had the highest and lowest average rating for Question 1 (I learnt a lot during the course.) across all their courses?**

```{r}
#| label: question-1-high-low
# code chunk for Q9

evals_clean |>
  filter(question_no == 901) |>
  arrange(SET_score_avg) |>
  select(teacher_id, SET_score_avg) |>
  slice_min(SET_score_avg) # looking at instructors with the lowest SET score average

evals_clean |>
  filter(question_no == 901) |>
  arrange(SET_score_avg) |>
  select(teacher_id, SET_score_avg) |>
  slice_max(SET_score_avg) # looking at instructors with the highest SET score average

```

> The instructors that have the lowest SET score average for question 1 are 	100132, 3577,	40937, 37035,	54201, and 37011 while the instructores there are 361 teachers that have the highest SET score average for quesstion 1(ex: 40903, 84689, 38238, 38335, etc).

**11. Which instructors with one year of experience had the highest and lowest average percentage of students failing in the current semester across all their courses?**

```{r}
#| label: one-year-experience-failing-students
# code chunk for Q10

evals_clean |>
  filter(seniority == 1) |>
  count(teacher_id, percent_failed_cur) |>
  select(teacher_id, percent_failed_cur) |>
  arrange(percent_failed_cur) |>
  slice_min(percent_failed_cur) # the instructors that have the lowest average percent of students failing in current semester

evals_clean |>
  filter(seniority == 1) |>
  count(teacher_id, percent_failed_cur) |>
  select(teacher_id, percent_failed_cur) |>
  arrange(percent_failed_cur) |>
  slice_max(percent_failed_cur) # the instructors that have the highest average percent of students failing in current semester

```

> There are 9 instructors with the lowest average percentage of student's failing in the currect semester and they are 80974, 84688, 84689, 86222, 98650, 98651, 102379, 103092, and 106126. There is one instructure with the highest average percentage of student's failing in the current semester and they are 104362.

**12. Which female instructors with either a doctorate or professional degree had the highest and lowest average percent of students responding to the evaluation across all their courses?**

```{r}
#| label: female-instructor-student-response
# code chunk for Q11

evals_clean |>
  filter(sex == "female" & academic_degree == "dr" | sex == "female" & academic_degree == "prof") |>
  arrange(resp_share) |>
  count(teacher_id, resp_share) |>
  select(teacher_id, resp_share) |>
  slice_min(resp_share) # looking at which female instructor who's either a dr or prof degree that has the lowest average in student responding

evals_clean |>
  filter(sex == "female" & academic_degree == "dr" | sex == "female" & academic_degree == "prof") |>
  arrange(resp_share) |>
  count(teacher_id, resp_share) |>
  select(teacher_id, resp_share) |>
  slice_max(resp_share) # looking at which female instructor who's either a dr or prof degree that has the highest average in student responding

```

> The female instructor with either a doctorate or professor degree that has the lowest average percent of student's responding is 3485 while the female instructor with the highest average percentage of student's repsonding with the same conditions is 101508.